Extended Data

### 2022-05-22 doc/manuscript/Dow_manuscript_Nature.Rmd
Add RZ to acknowledgements
cut abstract 201-->200 words

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[<!--@jeong_phenology_2011;-->@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020<!--@stuble_plant_2021;@delgado_differences_2020-->] and commencement of net carbon dioxide (CO~2~) uptake[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[<!--@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;--><!--@jeong_phenology_2011;-->@finzi_carbon_2020<!--@stuble_plant_2021-->] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;<!--@crabbe_extreme_2016;-->@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect tree stem growth[@savage_coordination_2021;@delpierre_temperate_2016], which sequesters carbon (C) in wood that has a long residence time in the ecosystem[@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two forests, we show that warmer spring temperatures shifted stem diameter growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual ring width was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that any extra CO~2~ uptake in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011<!--;@crabbe_extreme_2016-->] does not significantly contribute to increased sequestration in long-lived woody stem biomass.**
**Rather, contradicting projections from global C cycle models[@richardson_terrestrial_2012;@keenan_net_2014], our empirical results imply that warming spring temperatures are unlikely to increase woody productivity enough to strengthen the long-term CO~2~ sink of temperate deciduous forests.**

@@

# Acknowledgements {-} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig, members of the ForestGEO Ecosystems & Climate Lab at SCBI, and four anonymous reviewers for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.

>>>>

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[<!--@jeong_phenology_2011;-->@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020<!--@stuble_plant_2021;@delgado_differences_2020-->] and commencement of carbon dioxide (CO~2~) uptake[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[<!--@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;--><!--@jeong_phenology_2011;-->@finzi_carbon_2020<!--@stuble_plant_2021-->] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;<!--@crabbe_extreme_2016;-->@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect tree stem growth[@savage_coordination_2021;@delpierre_temperate_2016], which sequesters carbon (C) in wood that has a long residence time in the ecosystem[@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two forests, we show that warmer spring temperatures shifted stem diameter growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual ring width was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that any extra CO~2~ uptake in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011<!--;@crabbe_extreme_2016-->] does not significantly contribute to increased sequestration in long-lived woody stem biomass.**
**Rather, contradicting projections from global C cycle models[@richardson_terrestrial_2012;@keenan_net_2014], our empirical results imply that warming spring temperatures are unlikely to increase woody productivity enough to strengthen the long-term CO~2~ sink of temperate deciduous forests.**

@@

# Acknowledgements {-} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig, Roman Zweifel, members of the ForestGEO Ecosystems & Climate Lab at SCBI, and three anonymous reviewers for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.
